---
title: "Iteration and List Columns"
author: "Jenna Mohammed"
date: '2023-11-08'
output: html_document
---

### For Loops 
- output object (something to save), iterate over some sequence. inside will have some body of code, input object as well
 
 format 
 - input list
 - output list
 - index that keeps track of what iteration that you're on
 
### loop functions 

loops can get to be a bit tedious, loop functions are a way to clean up loops 

idea: give me an input list and function, i will give an output

purrr:map()

### map 

goal of map is to clarify loop process 

will provide an output. focusing on input, function, output

output = map(input, f)


### lists 

can be considered a dataframe

powerful tool for iteration

can map a function to each element of a input list (dataframe), export the output list, and save it in the same dataframe. Keeps everything together in one dataframe (easier to keep track of)

```{r}
library(tidyverse)
library(rvest)
```

## lists 
 you can put anything ina list 
 
```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Jenns")
vec_logical = c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE)
mat = matrix(1:8, nrow = 2, ncol = 4)

summary = summary(rnorm(100))

# if you'd want to put these into a dataframe, you need to make sure you have the same number of elements in each of these

# you can't put all of this into a dataframe, but you can put it all into a list 

l = list(
  vec_numeric = 5:8,
  vec_char = c("My", "name", "is", "Jenns"),
  vec_logical = c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE),
  mat = matrix(1:8, nrow = 2, ncol = 4),
  summary = summary(rnorm(100))
  
)
  
print(l)

# the below 3 do the same thing 

l$vec_numeric
l[[1]]
l[["vec_numeric"]]

```
 
 ## `for` loop
 
```{r}
list_norm = 
  list(
    a = rnorm(20, mean = 3, sd = 1),
    b = rnorm(30, mean = 0, sd = 5),
    c = rnorm(40, mean = 10, sd = 2),
    d = rnorm(20, mean = -3, sd = 1))
    #^ 4 lists, each a sample from a normal distribution
    
```
 

```{r}
list_norm 
```
 
 pause and get my old function
 
 