---
title: "Writing Functions"
author: "Jenna Mohammed"
date: '2023-11-06'
output: html_document
---

# Iteration
 Iteration - the process of doing the same thing over and over again
 
### Arguments 
 
- what going into the function (ex. x in mean(x)
- can take default values
- use readable names

### Body

- Data input checks using conditional execution 
- perform operations 
- format output

### Return -> this specific thing

### Scoping 
- variables in global enviornment, when you execute a function it creates an environment 
- problematic if function is using a vairable that exists somewhere in your code but not in the global enviornment
- you want to make sure that the code inside of your function is entirely self contained

### conditional execution 

### how to write functions 

- start very small , avoid unneeded complexity


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

## Do something simple 

If i wanted to compute z scores, i need to take whatever my input vector is and take the mean and sd of it and do the operation 

```{r}
# take sample and convert into collection of z scored 

x_vec = rnorm(30, mean = 5, sd = 3)

# will take the z score 

(x_vec - mean(x_vec)) / sd(x_vec)

```

I want a function to compute z score 

```{r}
z_scores = function(x) {

 z =  (x - mean(x)) / sd(x)  
 
 return(z)
  
}
  
# everything that happens between the brackets happens inside the body of the function

# we want the body of the function to depend on "x" since that is what it is called -> only want the variable x 

# we have written the function ^ and it is waiting for us to operate on it 

z_scores(x_vec)

# we are telling z scores that the input ^ that whenever we see an x in the body of a function, that is where we are going to plug in "x_vec
```

Try my function on some other things. These should have errors 

```{r error=TRUE}
z_scores(3)

# ^ cannpt compute sd of 3 so it will give you Na

z_Scores("my name is Jenna")

# ^ cannot take mean of a character so it won't work

z_scores(mtcars)

# cannot take mean of a dataset so it won't work

z_scores(c(TRUE, TRUE, FALSE, TRUE))

# this works, reads this as a sequence and turns it into 0's and 1's , converts to 1101 and can then compute the mean and sd 
```

let's update the function 

goal -> if someone tries to input something that doesn't make sense we want this to not work 

```{r}
z_scores = function(x) {
  
 if (!is.numeric(x)) {
   stop("Input must be numeric")
 }

  if (length(x) < 3) {
    stop("Input must have at least 3 numbers")
  }

 z =  (x - mean(x)) / sd(x)  
 
 return(z)
  
}

# if someone passes in something that is logical, the function is going to be false

# the "Stop" function allows you to express a message when the function doesn't work

z_scores(x_vec)

```

Try y funtion on some other things (again), these should have errors 
```{r error=TRUE}
z_scores(3)

z_Scores("my name is Jenna")

z_scores(mtcars)

z_scores(c(TRUE, TRUE, FALSE, TRUE))

```

## Multiple Outputs 

making it more complex 

compute mean and sd at the same time 

```{r}
mean_and_sd = function(x) {
  
 if (!is.numeric(x)) {
   stop("Input must be numeric")
 }

  if (length(x) < 3) {
    stop("Input must have at least 3 numbers")
  }

 mean_x = mean(x) 
 sd_x = sd(x)
 
 list(
   mean = mean_x,
   sd = sd_x
 )
 
 # if you have a variety of things that you want to put toether, you can make that a list
 
 tibble(
   mean = mean_x,
   sd = sd_x
 )
  
 # ^ making this a dataframe (in a table)
}
```

check that the function works 

```{r}
mean_and_sd(x_vec)
```

we passed in an input vector x, and got 5.54 as the mean and 2.67 as the sd 

## Multiple Inputs 

This function will say "give me a sample size, mean, and sd and I am going to spit things out"

Up until this point we have been creating a vector and spitting it out


I'd like to do this with a function 

```{r}
# create a simulated dataset

sim_data = 
  tibble(
    x = rnorm(n = 100, mean = 4, sd = 3)
  )

# we have 3 inputs, the sample size, mean, and sd

sim_data |>
  summarize(
  mean = mean(x), 
  sd = sd(x)
)
```


```{r}
# for the function, we'll say that we need an input sample size, an input mean, and an input sd 

sim_mean_sd = function(sample_size, mu, sigma) {
  
  sim_data = 
  tibble(
    x = rnorm(n = sample_size, mean = mu, sd = sigma)
  )


sim_data |>
  summarize(
    mean = mean(x), 
    sd = sd(x)
)
}

# 1st argument = 1st argument, and so on
sim_mean_sd(100,6,3)

# or named matching (typically should use this(could be in different order))

sim_mean_sd(sample_size = 100,mu = 6, sigma = 3)

# can also make default values for the original function

#"sim_mean_sd = function(sample_size, mu = 3, sigma = 4)"

# we want it to show a table of about these numbers

# the function creates a dataframe based on the inputs, summarizes it to compute a mean and sd 



```

## the point of this IS

if i run this multiple times, each time is going to create a different normal distribution, mean, and sd. If i do this over and over again, then each time i will get a different mean and different sd, and over time i can learn about that process. how much we might expect the mean to shift away from the mean value

In general, if i am able to do this again and again and again, i would start to be able to get some idea of what the mean and sd look like incomparison to the true value

