---
title: "Writing Functions"
author: "Jenna Mohammed"
date: '2023-11-06'
output: html_document
---

# Iteration
 Iteration - the process of doing the same thing over and over again
 
### Arguments 
 
- what going into the function (ex. x in mean(x)
- can take default values
- use readable names

### Body

- Data input checks using conditional execution 
- perform operations 
- format output

### Return -> this specific thing

### Scoping 
- variables in global enviornment, when you execute a function it creates an environment 
- problematic if function is using a vairable that exists somewhere in your code but not in the global enviornment
- you want to make sure that the code inside of your function is entirely self contained

### conditional execution 

### how to write functions 

- start very small , avoid unneeded complexity


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

## Do something simple 

If i wanted to compute z scores, i need to take whatever my input vector is and take the mean and sd of it and do the operation 

```{r}
# take sample and convert into collection of z scored 

x_vec = rnorm(30, mean = 5, sd = 3)

# will take the z score 

(x_vec - mean(x_vec)) / sd(x_vec)

```

I want a function to compute z score 

```{r}
z_scores = function(x) {

 z =  (x - mean(x)) / sd(x)  
 
 return(z)
  
}
  
# everything that happens between the brackets happens inside the body of the function

# we want the body of the function to depend on "x" since that is what it is called -> only want the variable x 

# we have written the function ^ and it is waiting for us to operate on it 

z_scores(x_vec)

# we are telling z scores that the input ^ that whenever we see an x in the body of a function, that is where we are going to plug in "x_vec
```

Try my function on some other things 

```{r error=TRUE}
z_scores(3)

# ^ cannpt compute sd of 3 so it will give you Na

z_Scores("my name is Jenna")

# ^ cannot take mean of a character so it won't work

z_scores(mtcars)

# cannot take mean of a dataset so it won't work

z_scores(c(TRUE, TRUE, FALSE, TRUE))

# this works, reads this as a sequence and turns it into 0's and 1's , converts to 1101 and can then compute the mean and sd 
```

let's update the function 

goal -> if someone tries to input something that doesn't make sense we want this to not work 

```{r}
z_scores = function(x) {
  
 if (!is.numeric(x)) {
   stop("Input must be numeric")
 }

  if (length(x) < 3) {
    stop("Input must have at least 3 numbers")
  }

 z =  (x - mean(x)) / sd(x)  
 
 return(z)
  
}

# if someone passes in something that is logical, the function is going to be false

# the "Stop" function allows you to express a message when the function doesn't work

z_scores(x_vec)

```

```{r error=TRUE}
z_scores(3)

z_Scores("my name is Jenna")

z_scores(mtcars)

z_scores(c(TRUE, TRUE, FALSE, TRUE))

```
